Q1: Are lake and terrestrial primary productivity coherent?

### 2019-05-30 AnalysisCode/ms1_analyses.Rmd
add maps to analyses

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)
library(lmerTest)

@@

@@

#agriculture -- is 500m buffer best? Other options include 100m buffer (probably too small) and hu12 watershed
pct.ag<-lagosne_select(table="buffer500m.lulc", vars=c("lagoslakeid","buffer500m_nlcd2001_pct_82","buffer500m_nlcd2006_pct_82","buffer500m_nlcd2011_pct_82"))
pct.ag<-pct.ag[pct.ag$lagoslakeid %in% analysislakes$lakeinfo$lagoslakeid,]
pct.ag.avg<-data.frame(lagoslakeid=pct.ag$lagoslakeid, pct.ag=rowMeans(pct.ag[,2:4]))

@@

```{r gls modelling, echo=TRUE, cache=FALSE}
#short timescales
gls.coh.accndvi.st<-gls(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.st)
dredge.coh.accndvi.st<-dredge(gls.coh.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

gls.p.accndvi.st<-gls(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.p.accndvi.st)
dredge.p.accndvi.st<-dredge(gls.p.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

gls.phi.accndvi.st<-gls(cos(accndviphi.ts1) ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi.phist) #remove ftype b/c only lakes
summary(gls.phi.accndvi.st)
dredge.phi.accndvi.st<-dredge(gls.phi.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.st))

#long timescales
gls.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.lt)
dredge.coh.accndvi.lt<-dredge(gls.coh.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

gls.p.accndvi.lt<-gls(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.p.accndvi.lt)
dredge.p.accndvi.lt<-dredge(gls.p.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

gls.phi.accndvi.lt<-gls(cos(accndviphi.ts2) ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi.philt,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.phi.accndvi.lt)
dredge.phi.accndvi.lt<-dredge(gls.phi.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.lt))
```

@@

```{r mixed modelling, echo=TRUE, cache=FALSE}
lme.coh.accndvi.st<-lmer(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.coh.accndvi.st)
dredge.coh.accndvi.st<-dredge(lme.coh.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

lme.p.accndvi.st<-lmer(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.p.accndvi.st)
dredge.p.accndvi.st<-dredge(lme.p.accndvi.st, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

lme.phi.accndvi.st<-lmer(accndviphi.ts1 ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi.phist, na.action="na.fail")
summary(lme.phi.accndvi.st)
dredge.phi.accndvi.st<-dredge(lme.phi.accndvi.st, beta="sd") #model with TSI is unconvincingly better than intercept-only (dAIC = 0.92)
print(head(dredge.phi.accndvi.st))

lme.coh.accndvi.lt<-lmer(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.coh.accndvi.lt)
dredge.coh.accndvi.lt<-dredge(lme.coh.accndvi.lt, beta="sd") #lake type is marginal improvement over intercept only (dAIC = 0.1)
print(head(dredge.coh.accndvi.lt))

lme.p.accndvi.lt<-lmer(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.p.accndvi.lt)
dredge.p.accndvi.lt<-dredge(lme.p.accndvi.lt, beta="sd") #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

lme.phi.accndvi.lt<-lmer(accndviphi.ts2 ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi.philt, na.action="na.fail")
summary(lme.phi.accndvi.lt)
dredge.phi.accndvi.lt<-dredge(lme.phi.accndvi.lt, beta="sd") #model with TSI is unconvincingly better than intercept-only (dAIC = 0.92)
print(head(dredge.phi.accndvi.lt))

>>>>

library(aqts)
library(wsyn)
library(rgdal)
library(LAGOSNE)
library(raster)
library(MuMIn)
library(dplyr)
library(nlme)
library(lmerTest)
library(viridis)
library(fields)

@@

```{r mapping, echo=TRUE, cache=FALSE}

states<-readOGR("~/Box Sync/NSF EAGER Synchrony/Data/statesp020.shp")
getstates<-c("Minnesota", "Iowa", "Wisconsin", "Illinois", "Missouri", "Michigan", "Indiana", "Ohio", "Pennsylvania", "New York", "New Jersey", "Connecticut", "New Hampshire", "Rhode Island", "Massachusetts", "Vermont", "Maine")
lagosstates<-states[states@data$STATE %in% getstates,]

plot(lagosstates, main="Lakes selected for analysis")
points(analysislakes$lakeinfo$nhd_long, analysislakes$lakeinfo$nhd_lat, pch=16, cex=1, col="blue")

cohplotdata<-left_join(analysislakes$lakeinfo, coh.chlaXaccndvi, by="lagoslakeid")

pal<-viridis(100)

par(mar=c(1,0,2,0))

plot(lagosstates, main="Lakes by short timescale coherence")
points(cohplotdata$nhd_long, cohplotdata$nhd_lat, pch=16, cex=1, col=pal[round(cohplotdata$accndvicoh.ts1,2)*100])
colorbar.plot(x=mean(par("usr")[1:2]),y=par("usr")[3],strip=1:100,col=pal,horizontal = T)

plot(lagosstates, main="Lakes by long timescale coherence")
points(cohplotdata$nhd_long, cohplotdata$nhd_lat, pch=16, cex=1, col=pal[round(cohplotdata$accndvicoh.ts2,2)*100])
colorbar.plot(x=mean(par("usr")[1:2]),y=par("usr")[3],strip=1:100,col=pal,horizontal = T)

@@

#agriculture -- is 500m buffer best? Other options include 100m buffer (probably too small) and hu12 watershed
pct.ag<-lagosne_select(table="buffer500m.lulc", vars=c("lagoslakeid","buffer500m_nlcd2001_pct_82","buffer500m_nlcd2006_pct_82","buffer500m_nlcd2011_pct_82"))
# pct.ag<-lagosne_select(table="hu12.lulc", vars=c("lagoslakeid","hu12_nlcd2001_pct_82","hu12_nlcd2006_pct_82","hu12_nlcd2011_pct_82"))
pct.ag<-pct.ag[pct.ag$lagoslakeid %in% analysislakes$lakeinfo$lagoslakeid,]
pct.ag.avg<-data.frame(lagoslakeid=pct.ag$lagoslakeid, pct.ag=rowMeans(pct.ag[,2:4]))

@@

```{r gls modelling, echo=TRUE, cache=FALSE}
#short timescales
gls.coh.accndvi.st<-gls(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.st)
suppressWarnings(dredge.coh.accndvi.st<-dredge(gls.coh.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

gls.p.accndvi.st<-gls(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.p.accndvi.st)
suppressWarnings(dredge.p.accndvi.st<-dredge(gls.p.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

gls.phi.accndvi.st<-gls(cos(accndviphi.ts1) ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi.phist) #remove ftype b/c only lakes
summary(gls.phi.accndvi.st)
suppressWarnings(dredge.phi.accndvi.st<-dredge(gls.phi.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.st))

#long timescales
gls.coh.accndvi.lt<-gls(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.coh.accndvi.lt)
suppressWarnings(dredge.coh.accndvi.lt<-dredge(gls.coh.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.lt))

gls.p.accndvi.lt<-gls(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.p.accndvi.lt)
suppressWarnings(dredge.p.accndvi.lt<-dredge(gls.p.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

gls.phi.accndvi.lt<-gls(cos(accndviphi.ts2) ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + huc2_code, data=modvars.accndvi.philt,
                       correlation=corExp(form = ~ nhd_lat + nhd_long))
summary(gls.phi.accndvi.lt)
suppressWarnings(dredge.phi.accndvi.lt<-dredge(gls.phi.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.phi.accndvi.lt))
```

@@

```{r mixed modelling, echo=TRUE, cache=FALSE}
lme.coh.accndvi.st<-lmer(accndvicoh.ts1 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.coh.accndvi.st)
suppressWarnings(dredge.coh.accndvi.st<-dredge(lme.coh.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.coh.accndvi.st))

lme.p.accndvi.st<-lmer(accndvip.ts1 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.p.accndvi.st)
suppressWarnings(dredge.p.accndvi.st<-dredge(lme.p.accndvi.st, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.st))

lme.phi.accndvi.st<-lmer(accndviphi.ts1 ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi.phist, na.action="na.fail")
summary(lme.phi.accndvi.st)
suppressWarnings(dredge.phi.accndvi.st<-dredge(lme.phi.accndvi.st, beta="sd")) #model with TSI is unconvincingly better than intercept-only (dAIC = 0.92)
print(head(dredge.phi.accndvi.st))

lme.coh.accndvi.lt<-lmer(accndvicoh.ts2 ~ maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.coh.accndvi.lt)
suppressWarnings(dredge.coh.accndvi.lt<-dredge(lme.coh.accndvi.lt, beta="sd")) #lake type is marginal improvement over intercept only (dAIC = 0.1)
print(head(dredge.coh.accndvi.lt))

lme.p.accndvi.lt<-lmer(accndvip.ts2 ~ tslength + maxdepth + nhd_ftype + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi, na.action="na.fail")
summary(lme.p.accndvi.lt)
suppressWarnings(dredge.p.accndvi.lt<-dredge(lme.p.accndvi.lt, beta="sd")) #intercept only is best model. Disappointing.
print(head(dredge.p.accndvi.lt))

lme.phi.accndvi.lt<-lmer(accndviphi.ts2 ~ maxdepth + lake_area_ha + pct.ag + chla + tsi.cat + (1|huc2_code), data=modvars.accndvi.philt, na.action="na.fail")
summary(lme.phi.accndvi.lt)
suppressWarnings(dredge.phi.accndvi.lt<-dredge(lme.phi.accndvi.lt, beta="sd")) #model with TSI is unconvincingly better than intercept-only (dAIC = 0.92)
print(head(dredge.phi.accndvi.lt))
